% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file. 

% Note that since 2010, USENIX does not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the 
% usepackage command, below.

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix2019,epsfig,endnotes}
\usepackage{bm}
\usepackage{url}
\usepackage{amsmath}
\usepackage{subfig}
\usepackage{algorithm} 
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{amsfonts}
\usepackage{algpseudocode}
\begin{document}

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Perceived Quality driven VR Streaming}

%for single author (just remove % characters)
\author{
{\rm Your N.\ Here}\\
Your Institution
\and
{\rm Second Name}\\
Second Institution
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
} % end author

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
\thispagestyle{empty}


\subsection*{Abstract}
In recent years, Virtual Reality (VR) video streaming becomes popular quickly. How to provide high user-perceived quality with limited bandwidth becomes the biggest problem in VR video streaming. Traditional viewpoint-driven VR streaming makes the assumption that perceived quality is only related to viewpoint-object distance, so it simply allocates high bitrate to objects near user's viewpoint. However, In this paper we prove that user perceived quality is not only related to viewpoint-object distance, it is also related to background luminance, texture complexity, viewpoint moving speed and Depth-of-Field. With consideration of these insights, we can save x\% bandwidth while providing the same perceived quality.

To build our PQVRS, we build a model to measure perceived quality in VR display, and design an object-based tiling scheme to cut video into several tiles which can be independently encoded / decoded. Then we present our method to enable client-side perceived quality computation and optimization. We implement a prototype of PQVRS and compare its performance with state-of-art VR streaming schemes. Experimental results show that proposed PQVRS saves x\% bandwidth without decrease of perceived quality.


\input{sections/introduction}
\input{sections/motivation}
\input{sections/insights}
\input{sections/jNDmodel}
\input{sections/tiling}
\input{sections/adaptation}
\input{sections/implementation}


{\footnotesize \bibliographystyle{acm}
\bibliography{../common/bibliography}}

\end{document}