\documentclass[sigconf]{acmart}

\usepackage[english]{babel}
\usepackage{blindtext}
\usepackage{bm}
\usepackage{url}
\usepackage{amsmath}
\usepackage{subfig}
\usepackage{algorithm} 
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{amsfonts}
\usepackage{algpseudocode}

% Copyright
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

\settopmatter{printacmref=false, printccs=false, printfolios=true}

% DOI
\acmDOI{}

% ISBN
\acmISBN{}

%Conference
%\acmConference[Submitted for review to SIGCOMM]{}
%\acmYear{2018}
%\copyrightyear{}

%% {} with no args suppresses printing of the price
\acmPrice{}


\begin{document}
\title{Perceived Quality driven VR Streaming}

%\titlenote{Produces the permission block, and copyright information}
%\subtitle{Extended Abstract}

\author{Paper \# XXX, XXX pages}
% \author{Firstname Lastname}
% \authornote{Note}
% \orcid{1234-5678-9012}
% \affiliation{%
%   \institution{Affiliation}
%   \streetaddress{Address}
%   \city{City} 
%   \state{State} 
%   \postcode{Zipcode}
% }
% \email{email@domain.com}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{X.et al.}

\begin{abstract}
In recent years, Virtual Reality (VR) video streaming becomes popular quickly. How to provide high user-perceived quality with limited bandwidth becomes the biggest problem in VR video streaming. Traditional viewpoint-driven VR streaming makes the assumption that perceived quality is only related to viewpoint-object distance, so it simply allocates high bitrate to objects near user's viewpoint. However, In this paper we prove that user perceived quality is not only related to viewpoint-object distance, it is also related to background luminance, texture complexity, viewpoint moving speed and Depth-of-Field. With consideration of these insights, we can save x\% bandwidth while providing the same perceived quality.

To build our PQVRS, we build a model to measure perceived quality in VR display, and design an object-based tiling scheme to cut video into several tiles which can be independently encoded / decoded. Then we present our method to enable client-side perceived quality computation and optimization. We implement a prototype of PQVRS and compare its performance with state-of-art VR streaming schemes. Experimental results show that proposed PQVRS saves x\% bandwidth without decrease of perceived quality.
\end{abstract}

\maketitle

\input{sections/introduction}
\input{sections/motivation}
\input{sections/insights}
\input{sections/jNDmodel}
\input{sections/tiling}
\input{sections/adaptation}
\input{sections/implementation}
\input{sections/evaluation}
\input{sections/Relatedworks}
\input{sections/conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{reference}

\end{document}